Knowledge Representation for Service Robots: Supply and Demand

### 2012-05-12 krs-survey.tex
Filled most of the 'Role representation' section

This section proposes a more formal typology of desirable features for a such a
system. For each feature, we provide a short definition along with links to
relevant literature.

@@

@@

As an agent acting at human-like time scale and dealing with temporal concepts
(like actions), a robot may want to represent, and possibly to reason, about
time. Time representation is split into two distinct abilities: representing
time points (both in the past -- which is roughly equivalent to assignment of
timestamps to events the robot perceives -- and in the future), and
representing \emph{passing time} (durations, timespans) like in \emph{``the
eggs will be cooked in 10 min''}.

@@

We call a system that do not account for time (\ie that permanently lives in
present) \emph{atemporal}.

\paragraph{Representation of events and actions}

>>>>

This section proposes a more formal typology and a nomenclature of analysis
axis for a such a system. For each feature, we provide a short definition along
with links to relevant literature.

@@

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf B.1 Roles}
            child {node [taxon] {{\bf B.1.3}. Actions}}
            child {node [taxon] {{\bf B.1.2}. Time}}
            child {node [taxon] {{\bf B.1.1}. Space}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

Symbolic representation of space is a widely studied topic. In particular, a
large literature corpus is available on spatial ontologies.

Two main classes of spatial relations are usually represented at the symbolic
level: the topology of environments and the placement of physical entities.

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf B.1.1 Space}
            child {node [taxon] {{\bf B.1.1.2}. Placement}}
            child {node [taxon] {{\bf B.1.1.1}. Topology}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

...topology??

Symbolic representation of entities placement can be absolute or relative. The
relation \concept{isOn}, for example, leads to absolute statements: the
validity of the relation is independent of the nature of its subjects and
objects, and is also independent of the observer.

On the contrary, the relation \concept{nextTo} is relative, and depends on the
relative size of the subject and the object. Two houses distant of 2 meters
from each other can be considered as next to each other. Two ants separated by
2 meters are not next to each other.

The relation \concept{leftOf} is another example of relative spatial relation,
this time because it depends on the observer viewpoint.

\fxwarning{\concept{front}/\concept{back} -> entities have or not a concept of front/back.}

Choices must be made in the knowledge representation system to adequately
represent relative spatial relations. Options include the computation of such
relations only on-demand, when the context is know (which viewpoints, etc.) or
storage of these facts in different models, one per agent.

\fxwarning{Find references on that subject -> classification of spatial
relations. Wikipedia?}

@@

As an agent acting at human-like time scale and dealing with temporal concepts
(like actions), a robot may want to represent, and possibly to reason, about
time. Time representation is split into two distinct abilities: representing
time points (both in the past -- which is roughly equivalent to assignment of
timestamps to events the robot perceives -- and in the future), and
representing \emph{passing time} (situations, durations, timespans) like in
\emph{``the eggs will be cooked in 10 min''}.

We call \emph{situations} all temporal concepts with a non-zero duration,
whereas \emph{events} are durationless temporal concepts.

@@

We call a system that does not account for time (\ie that permanently lives in
present) \emph{atemporal}.

\paragraph{Actions}

Events and actions are two temporal concepts that are of particular importance to
robotic systems, as systems that perceive, react and perform in their environment.

While representation of events \fxwarning{revient à} to label a timestamp,
representation of actions are more complex since they semantically imply
interactions between concepts. From a taxonomy point of view, actions are a
particular type of event that normaly leads to a situation corresponding to the
action realization.

\emph{Thematic roles}~\cite{Gruber1965} (also found as \emph{semantic roles} or
\emph{theta roles} in the literature) allow to semantically qualify the
parameters of an action. The recipient of the action, the performer, the object
acted upon, the destination are some example of common thematic roles.

Table~\ref{table|theta-roles} presents a more comprehensive list of thematic
roles proposed by \cite{Aarts1997} (see~\cite{Gutierrez2001} for a comparison
of other sets of thematic roles present in the literature).

\begin{tabular}{lp{5cm}}

\toprule
       Role & Meaning \\
\midrule
      Agent & The \emph{doer} or instigator of the action denoted by the predicate. \\
    Patient & The \emph{undergoer} of the action or event denoted by the predicate. \\
      Theme & The entity that is moved by the action or event denoted by the predicate. \\
Experiencer & The living entity that experiences the action or event denoted by the predicate. \\
       Goal & The location or entity in the direction of which something moves. \\
Benefactive & The entity that benefits from the action or event denoted by the predicate. \\
     Source & The location or entity from which something moves. \\
 Instrument & The medium by which the action or event denoted by the predicate is carried out. \\
   Locative & The specification of the place where the action or event denoted by the predicate in situated. \\
\bottomrule

\end{tabular}
\end{center}
\caption{Thematic Roles~\cite{Aarts1997}.}
\label{table|theta-roles}
\end{table}

In the context of robotic, a \emph{task} represents the abstract model of an
action. Tasks usually specify at least the conditions required to performed the
action and the consequences of the realization of the action, and are central
to the decisional layers of the robot, in particular for the planning,
monitoring and execution control activities. Knowledge representation systems
may thus have specific mechanisms to represent them (and possible, to reason
about them, as presented in section~\ref{sect|planning}).
